Data Preparation

### 2023-03-25 x_data-crunching.qmd
getting towards completing arrival trajectory extraction

```{r}
# we have identified an arrival trajectory
trj_arr <- arr_trjs |> filter(UID %in% "0200cb-0") |> 
#trj_arr <- arr_trjs |> filter(UID %in% c("00b1f9-0")) |> 
# trim payload and add sequence id
  select(UID, TIME, ICAO24, LAT, LON, ALT_B)  |> 
  rowid_to_column("NODE_ID") |> 
  mutate(MST = NA_character_)

ldg_rwy <- trj_arr |> assign_landing_runway(.arp = arp) |> 
  inner_join(rwys, by = "REF")
my_thr  <- ldg_rwy |> select(UID, REF, LAT, LON, COMMENT) |> 
  cast_latlon_to_pts(.drop_coord = FALSE)

# check if we have a threshold overflight
ovr_thr <- identify_threshold_overhead(trj_arr, my_thr)
if(nrow(ovr_thr) > 0){
  message("THRESHOLD")
  eldt    <- ovr_thr |> filter(DIST_THR == min(DIST_THR)) 
  trj_arr[eldt$NODE_ID,"MST"] <- "ALDT"
}else{
  message("NEED TO WORK ON LANDING TIME")
}
```

@@

check for blended point
final_sfn |> activate("nodes") |> filter(is.na(TIME)) |> select(TIME, ICAO24, LAT, LON, ALT_B, geometry)

point is injected, need to fill details, e.g. TIME
then extract nodes ~ enriched sequence of final points (maybe use sequence number)

>>>>

```{r}
source("~/RProjects/paper-2023-ICNS/R/assign_threshold_time.R")

# we have identified an arrival trajectory
trj_arr <- arr_trjs |> filter(UID %in% "0200cb-0") |> 
#trj_arr <- arr_trjs |> filter(UID %in% c("00b1f9-0")) |> 
  assign_threshold_time(my_thr)
```

@@

PICK SOME ARRIVALS FOR GETTING CODE IN PLACE

```{r}
# derived from validation data set, binned per 15 mins per arrival runway
# 27L 06:30 - 0645
peak_wave <- c("BAW377", "EWG6VP", "BAW11J", "ETD89G", "AIC161", "AFR82FE", "BAW236","DLH3WU","VIR138M","EWG30N")

peak_trjs <- arr_trjs |> filter(FLTID %in% peak_wave)

viz_check_DAO(peak_trjs) + viz_check_vertical(peak_trjs |> filter(TIME > lubridate::ymd_hm("2019-08-01 06:20")))
```

```{r}
# run loop with group_modify()
# create a new UID for the looping as group_modify cannot accept UID in payload
# TODO - write functions independent of UID or handle UID outside calculation.
study_trjs <- peak_trjs |> 
  mutate(LUID = UID) |> group_by(LUID) |> 
  group_modify(.f = ~assign_threshold_time(.x,.my_thr = my_thr)) |> 
  ungroup() |> select(-LUID)
```

```{r}
ggplot(data = study_trjs) +
  geom_point(aes(x = TIME, y = ALT_B)) +
  geom_vline(xintercept = study_trjs$TIME[which(study_trjs$MST == "ALDT")])
```

first pair

```{r}
# "BAW377","EWG6VP","BAW11J"
inter_arrival <- study_trjs |> filter(MST == "ALDT") |> group_by(MST_REF) |> arrange(TIME, .by_group = TRUE) |> 
  mutate(
     FOLLOWER = lead(UID, default = NA_character_)
    ,SEP_TIME = difftime(lead(TIME, default = NA), TIME, unit = "sec")
    )
inter_arrival
```

```{r}
p1 <- inter_arrival |> ggplot() +
  geom_step(aes(x = TIME, y = SEP_TIME))
p2 <- inter_arrival |> ggplot() +
  geom_boxplot(aes(x = MST_REF, y = SEP_TIME))
p1 + p2
```

OK - FUCKED UP HERE 
NEED to catch improper trajectories that throw error ==> improve cleaning
wrap look in safely/possibly ... to take care of any other uncaught exception

PRESS ON WITH PIPELINE ... data cleaning is on track ... should work for paper

Do this for all of 1. August

```{r}
fn <- here::here("data","validation-EGLL-APDF-20190801.csv.gz")
egll_arrs_apdf <- read_csv(fn, show_col_types = FALSE)

# safe loop
safe_assign_threshold_time <- possibly(.f = assign_threshold_time, otherwise = "Error")

study_trjs <- arr_trjs |> 
  filter(FLTID %in% egll_arrs_apdf$FLTID) |> 
  mutate(LUID = UID) |> group_by(LUID) |> 
  group_modify(.f = ~safe_assign_threshold_time(.x,.my_thr = my_thr)) |> 
  ungroup() |> select(-LUID)
```
